The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Corrections after spell check
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@760 4525493e-7705-40b1-a816-d608a930855b

\begin{abstract}
Metadata plays a significant role in large modern enterprises, 
research experiments and digital libraries where it comes from many different 
sources and is distributed in a variety of digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
an information retrieval approach to non-relational data sources,
we can't do so for relational ones, where information is accessed via
a pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore metadata associated with petabytes of experimental data. 
We combine the simplicity of keyword-based search with the precision of RDMS
under the new system. The aggregated information is collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
we have performed a set of detailed, large scale tests the results of which 
we present in this paper.
\end{abstract}

@@

As was pointed out in \cite{Arms} the mixed content and 
mixed meta-data and meta-data consistency 
of the underlying services needs to be carefully
considered in the design of the system if it is to be a successful information 
discovery service.
Starting from this ground we are designed the DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use conditional operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. It conists of cache server,
analytics and mapping databases. The data in a cache can be shared
across multiple queries. The information retrieval is triggered
on demand basis by invoking appropriate data-service APIs. The
optional DAS robots (Unix daemons) can be used to pre-fetch
the most populary query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
For instance, we didn't require any transactions and data persistency in DAS;
a dynamically typing of stored meta-data objects was one of the requirements, etc.
Those arguments forced us to look for alternative IT solutions.
Through the careful analysis of available options, such as a file-based and in memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
arbitrary document structure storage, replication and fail-over features. 
We made our decision in favor of MongoDB, due to its support of dynamic queries, 
full indexes, including inner objects and embedded arrays,
and auto-sharding. Our preliminary benchmarks have shown that it can sustain
the desired load and size capable of handling our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

@@

\subsection{DAS core\label{DAS_core}}
The DAS core library is responsible for data retrieval on demand. 
It has the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
Each input query was recorded into the Analytics DB for further 
analysis, and was decomposed into a set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requestor
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementations are similar and their 
logic is largely abstracted, apart from a few extentions such as specific
security modules for data-services accessible via private network.
Finally, the results were retrieved from the cache and delivered to the user.

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB is used to track services known to DAS and how to interact with them.
%The Mapping DB represents DAS {\it registration} service. 
It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
Each key is mapped into set of API calls, whose output is retrieved,
cached and aggregated as required.

@@

\subsection{DAS caching system}
The DAS caching system is used to dynamically fetch
and aggregate data upon user requests. We use the following 
algorithm (code snippets are written in Python \cite{Python}):
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+).
\item Check cache for query hash. If found, look-up results from the cache,
otherwise proceed to the next step.
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+r+ represents a triplet \verb+url, api, args+ and 
\verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery the data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(daskey):
   existing_record = get_from_cache(daskey)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+daskey+
is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
We also have an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into the cache.

\subsection{DAS queries}
All DAS queries are expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
is based on the analysis of daily work operations of our users, see \cite{DBS07}.
Each key represents a DAS record. Every record is stored into the cache
in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
we are able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
set of attributes representing JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
is associated with a \verb+block+ key who has the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record and therefore
a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
are able to specify a context of their queries by using appropriate 
attribute(s). For example, query \verb+block+ represents search for all
block records rather then ``block'' string, while query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matched names. The numerical values, conditions, grouping are
easy to express too, e.g. \verb+run > 10 and run < 100+ query 
is used to get all run records for provided range of run numbers. 
To complement the strength of the queries, we do plan to index 
dynamically DAS records and add a keyword-based 
search within record content later.

All input queries are transformed into MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run>10+
is represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries are stored into Analytics DB similar to other DAS 
records and used for further analysis.

\subsection{Data-services and aggregation}
As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provide information about CMS blocks. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas.

Upon a user request to 
find a block meta-data, we were able to identify appropriate set of 
DBS and PhEDEx APIs, retrieve data, re-format their output 
into DAS notations and store results into DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such an aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ refers to api call records stored in Analytics DB,
the \verb+block+ part represents an aggregation block meta-data information from DBS
and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS data location and file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system is divided into dozen 
of individual services (instances) with own size and lifetime 
used by different production teams for various bookkeeping tasks; 
the RunSummary data-service is operated on private network accessible 
at CMS online operation center.

At the moment, the amount of meta-data stored in all services is around
75 GB and 200 GB in tables and index sizes, respectively. We estimate to
collect $\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we ere measured elapsed time and CPU/RAM utilization
required to fetch, store and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

@@

The initial DAS implementation is done in Python language \cite{Python}
and all tests are performed on x64 Linux platform where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
All tests are performed multiple times to avoid fluctuations. 
The elapsed time is measured and consists of the following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here {\it retrieval time} is a time required to access data from data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert data-service meta-data to DAS records for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
cache back-end, {\it aggregation time} is required to merge objects into DAS records based
on their common key (\verb+block.name+ in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, are only applied to
one of the sub-system, either DBS or PhEDEx, depending on the insertion order.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component is bigger 
due to longer retrieval and parsing time as well as time spent
during aggregation step. As shown in Table \ref{DAS_benchmark},
the look-up time spent to access individual components (last column) is quite
reasonable, roughly 1 second for both systems. While the final time to
get DAS records on disk is about 50 seconds. This is mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on the results of stress tests we are measured that we achieve throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for 
reading operations in DAS, which is suitable for our needs. Since all 
operations are performed on a single CPU, a further improvements are expected. 
Also, a necessary tuning can be done by using C/C++ driver instead of python one.

\section{Summary}
We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All data are retrieved on demand basis,
while pre-fetching of most common queries is in our plans. We are developed a
prototype in Python which is under commissioning phase right now.
The performance studies of the DAS system are presented in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced each year by experiment will be of the order of
500GB. The DAS system should sustain such load and provide generic 
``data-discovery'' service for CMS experiment in years to come.

@@

This work was supported by the National Science Foundation and 
Department of Energy of the United States of America. 
Fermilab is operated by Fermi Research Alliance, LLC under Contract
No. DE-AC02-07CH11359 with the United States Department of Energy.

>>>>

\begin{abstract}
Meta-data plays a significant role in large modern enterprises, 
research experiments and digital libraries where it comes from many different 
sources and is distributed in a variety of digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
an information retrieval approach to non-relational data sources,
we can't do so for relational ones, where information is accessed via
a pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine the simplicity of keyword-based search with the precision of RDMS
under the new system. The aggregated information is collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
we have performed a set of detailed, large scale tests the results of which 
we present in this paper.
\end{abstract}

@@

As was pointed out in \cite{Arms} the mixed content and 
mixed meta-data and meta-data consistency 
of the underlying services needs to be carefully
considered in the design of the system if it is to be a successful information 
discovery service.
Starting from this ground we are designed the DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use conditional operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. It consists of cache server,
analytics and mapping databases. The data in a cache can be shared
across multiple queries. The information retrieval is triggered
on demand basis by invoking appropriate data-service APIs. The
optional DAS robots (Unix daemons) can be used to pre-fetch
the most popular query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
For instance, we didn't require any transactions and data persistency in DAS;
a dynamically typing of stored meta-data objects was one of the requirements, etc.
Those arguments forced us to look for alternative IT solutions.
Through the careful analysis of available options, such as a file-based and in memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems are ``schema-less'' databases providing
arbitrary document structure storage, replication and fail-over features. 
We made our decision in favor of MongoDB, due to its support of dynamic queries, 
full indexes, including inner objects and embedded arrays,
and auto-sharding. Our preliminary benchmarks have shown that it can sustain
the desired load and size capable of handling our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

@@

\subsection{DAS core\label{DAS_core}}
The DAS core library is responsible for data retrieval on demand. 
It has the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
Each input query was recorded into the Analytics DB for further 
analysis, and was decomposed into a set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requester
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementations are similar and their 
logic is largely abstracted, apart from a few extensions such as specific
security modules for data-services accessible via private network.
Finally, the results were retrieved from the cache and delivered to the user.

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB is used to track services known to DAS and how DAS should interact with them.
It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
Each key is mapped onto set of API calls, whose output is retrieved,
cached and aggregated as required.

@@

\subsection{DAS caching system}
The DAS caching system is used to dynamically fetch
and aggregate data upon user requests. We use the following 
algorithm (code snippets are written in Python \cite{Python}):
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+).
\item Check cache for query hash. If found, look-up results from the cache,
otherwise proceed to the next step.
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+r+ represents a triplet \verb+url, api, args+ and 
\verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery the data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(daskey):
   existing_record = get_from_cache(daskey)
   merge(row, existing_record)
   insert_into_cache(row)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+daskey+
is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
%We also have an option to run independent set of daemons, DAS robots, who
%monitor Analytics DB for most popular requests and pre-fetch the most
%common data into the cache. 
We can also automate populating the cache by running an 
independent set of daemons, DAS robots, which monitor the 
Analytics DB for suitable requests (for example queries that are 
popular, periodic in nature or expensive) and pre-fetch data into the cache.

\subsection{DAS queries}
All DAS queries are expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. The choice of keys
is based on the analysis of day to day user operations \cite{DBS07}.
Each key represents a DAS record. Every record is stored into the cache
in JSON format. Due to the ``schema-less'' nature of the underlying MongoDB
back-end we are able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
set of attributes describing its JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
is associated with a \verb+block+ key who has the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record, allowing
for nested structures, e.g. \verb+block.replica.size+. Our users 
are able to specify the context of their queries by using appropriate 
attribute(s). For example, a query for \verb+block+ represents a search for all
block records rather then the ``block'' string, while the query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matching names. Numerical values, conditions and grouping are
also easy to express, for example, a query
of the form \verb+run > 10 and run < 100+ can be used to
retrieve all run records for the given range of run numbers. 
As a complement to the queries, we will also be adding a dynamic index
of DAS records with a keyword-based search within record content later.

All input queries are transformed into the MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run > 10+
is represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries are stored into Analytics DB similar to other DAS 
records and used for further analysis.

\subsection{Data-services and aggregation}
As discussed in section \ref{MappingDB} the DAS Mapping DB 
is the authoritative source of information on the relationships between 
all the data services known to DAS.
%became an authorative
%source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we are able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provide information about CMS file block structure and locations.
DBS associates groups of files into blocks, while PhEDEx tracks the blocks
location. 
%In the former case, DBS 
%stored information about blocks and their relationships to other 
%physics objects, such as dataset and files, while PhEDEx kept
%information about its location and replicas.

When a query for block meta-data is made, we are
%Upon a user request to 
%find a block meta-data, we were 
able to identify the appropriate set of 
DBS and PhEDEx APIs, retrieve the raw data from the services, format it
for DAS use and add the resulting information to the DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such an aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ indicates the records stored in the DAS Analytics DB,
the \verb+block+ part represents the aggregated block meta-data information from DBS
and PhEDEx and the \verb+das+ part contains expiration time stamps
about how long the information is valid.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS data location and file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiarities. 
For instance, the data in PhEDEx is transient due to constant 
migration of CMS data; the DBS system is divided into several individual 
instances with sizes and lifetimes driven by specific processing tasks; 
the RunSummary data-service is operated on private network accessible 
at CMS online operation center.

At the moment, the amount of meta-data stored in all services is around
75 GB and 200 GB in tables and index sizes, respectively. We estimate to
collect around 500 GB of meta-data each year during 
normal CMS operations. The average query rate to DAS is expected to be at the 
$\leq$10K requests a day.
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured the elapsed time and CPU/RAM utilization
required to fetch, store and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

@@

The initial DAS implementation is done in Python language \cite{Python}.
All tests are performed on a 64 bit Linux platform with
1 CPU for the DAS server and 1 CPU for the MongoDB database server. 
All tests are performed multiple times to avoid fluctuations. 
The elapsed time is measured and consists of the following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here {\it retrieval time} is the time required to access data from data-service,
{\it parsing time} is the time required to read and parse the received data
from the services, and {\it re-mapping time} is the time required to convert 
that information into DAS. {\it Cache insertion} and {\it indexing time} 
represents time spent on DAS back-end caching, {\it aggregation time} is
the time required to merge objects into DAS records based
on their common keys (\verb+block.name+ in this case) and {\it output creation time}
is the time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, are only applied to
one of the sub-system, either DBS or PhEDEx, depending on the insertion order.

@@

The individual tests of DAS components (DBS and PhEDEx) show roughly 
the same performance, although the time taken by the PhEDEx component is longer
due to higher retrieval and parsing time as well as time spent
during aggregation step. As shown in Table \ref{DAS_benchmark},
the look-up time spent to access individual components (last column) is quite
reasonable, roughly 1 second for both systems. While the final time to
get DAS records on disk is about 50 seconds. This is mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

The results of the stress tests have shown that we can achieve a throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
needs. Please note, we have relatively small audience of users
who perform complex queries, while search engines designed for
a large number of users asking for imprecise, simple data.
Since all tests were performed on a single CPU, further improvements from
expanding to multiple CPUs are expected. In addition, more
performance advantage can be gained by writing dedicated C/C++
python extensions for data intensive operations.

\section{Summary}
%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
%aggregation layer on top of the existing relational and non-relation data-services
%mostly in real time fashion. All data are retrieved on demand basis,
%while pre-fetching of most common queries is in our plans. We are developed a
%prototype in Python which is under commissioning phase right now.
%The performance studies of the DAS system are presented in this paper. 
%The CMS experiment is started to collect its data in December of 2009 and 
%we expect to accumulate a few PB of real data each year. 
%In addition, the Monte-Carlo samples will be produced at this scale.
%Based on test studies performed in CMS, we expect that total size of
%meta-data produced each year by experiment will be of the order of
%500GB. The DAS system should sustain such load and provide generic 
%``data-discovery'' service for CMS experiment in years to come.
In this paper we have presented a new data aggregation service (DAS) 
developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
We have developed a prototype for DAS, which has been designed to provide 
caching and aggregation layer on top of the existing relational and 
non-relational data-services in a close to real time fashion. The data 
is retrieved on demand, with pre-fetching of common queries, 
determined from an extensive analytics database, also possible. 
The DAS system is currently being commissioned for production use by CMS.

Data taking at CMS began in December 2009, and, alongside simulation 
data, we expect to accumulate petabytes of data yearly. This will have 
of the order of a terabyte of associated meta-data that will need to be 
easily queried by physicists. From the performance studies of DAS presented 
in this paper we expect that the system will be able to sustain such load 
and will provide generic “data-discovery” service for CMS experiment for 
years to come.

@@

This work was supported by the National Science Foundation, 
contract No. PHY-0757894, and 
Department of Energy of the United States of America.

%Fermilab is operated by Fermi Research Alliance, LLC under Contract
%No. DE-AC02-07CH11359 with the United States Department of Energy.
